---
title: "Inequality in Happiness by Population Subgroups"
output: html_document

knit: (function(input_file, encoding) {
  out_dir <- '_site';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'inequality_analysis.html'))})
---


#### Define Libraries
```{r echo=TRUE, message=FALSE, warning=FALSE, tidy=TRUE}
library("stringr")
library("dplyr")
library("dineq")
library("DescTools")
library("ggplot2")
library("gridExtra")
```

#### Define Path
```{r tidy=TRUE}
dir.wrk <- getwd()
dir.data <- file.path(dir.wrk, "dataset/data")
dir.annot <- file.path(dir.wrk, "dataset/annotation")
```
#### Define Files
```{r tidy=TRUE}
file.dat <- file.path(dir.data, "mastertbl_household_reconstruction.tsv")
```

#### Load Data
```{r tidy=TRUE}
dat <- read.delim(file.dat, header=TRUE, stringsAsFactors=FALSE)

# FACTORIZE ---
dat$DISTRICT <- factor(dat$DISTRICT, levels=c("Dhading","Sindhupalchowk"))
dat$VDC_MUN <- factor(dat$VDC_MUN, levels=c("Gumdi","Sertung","Batase","Pangtang"))
dat$WARD <- as.factor(dat$WARD)
dat$SEX <- factor(dat$SEX, levels=c("Male","Female"))
dat$ETHNICITY <- factor(dat$ETHNICITY, levels=c("Bahun","Chhetri","Dalit-Others","Gurung-Magar","Newar","Tamang"))
dat$EDUCATION_LEVEL <- factor(dat$EDUCATION_LEVEL, levels=c("Illiterate","Primary","Secondary","University"))
dat$OCCUPATION <- factor(dat$OCCUPATION, levels=c("Agriculture","Business","Goverment Employee","Labour","Teacher"))
dat$INCOME_LEVEL <- factor(dat$INCOME_LEVEL, levels=c("0-2.5K","2.5-10K","10-20K","20-40K","40-60K","60K-Above"))
dat$RECON_COMPLETED <- as.factor(dat$RECON_COMPLETED)
#dat$SATISFACTION_LEVEL <- as.factor(dat$SATISFACTION_LEVEL)
dat$FREEDOM_CHOICE <- as.factor(dat$FREEDOM_CHOICE) 


#str(dat)
```
</br></br>

### Estimate Gini coefficient
```{r}
round(dineq::gini.wtd(x=dat$SATISFACTION_LEVEL, weights=NULL),3)

round(DescTools::Gini(x=dat$SATISFACTION_LEVEL, unbiased=TRUE, conf.level=NA, R=1000, type="bca", na.rm=FALSE),3)
```

</br></br>

### Decomposition by population subgroups (Mookharjee and Shorrocks, 1982)
</br>

#### *By Sex*
```{r}
sex_decomp <- dineq::gini_decomp(x=dat$SATISFACTION_LEVEL, z=dat$SEX)

df.sex_decomp <- data.frame(Categories=names(sex_decomp$gini_decomp), Value=round(as.numeric(sex_decomp$gini_decomp),3))

knitr::kable(df.sex_decomp, align="ll")
```

</br>

#### *By Ethnicity*
```{r}
eth_decomp <- dineq::gini_decomp(x=dat$SATISFACTION_LEVEL, z=dat$ETHNICITY)

df.eth_decomp <- data.frame(Categories=names(eth_decomp$gini_decomp), Value=round(as.numeric(eth_decomp$gini_decomp),3))

knitr::kable(df.eth_decomp, align="ll")
```

</br>

#### *By Income Group*
```{r}
inc_decomp <- dineq::gini_decomp(x=dat$SATISFACTION_LEVEL, z=dat$INCOME_LEVEL)

df.inc_decomp <- data.frame(Categories=names(inc_decomp$gini_decomp), Value=round(as.numeric(inc_decomp$gini_decomp),3))

knitr::kable(df.inc_decomp, align="ll")
```


</br></br>

### Calculate Gini of different groups

</br>

#### *By Sex*
```{r}
mean_satisfaction_level <- mean(dat$SATISFACTION_LEVEL, na.rm=TRUE)

df_sex1 <- dplyr::group_by(dat, SEX) %>%
                  dplyr::summarise(count = n(),
                            mean = round(mean(SATISFACTION_LEVEL, na.rm = TRUE), 3),
                            sd = round(sd(SATISFACTION_LEVEL, na.rm = TRUE), 3),
                            norm_mean = round(mean/mean_satisfaction_level, 3),
                            norm_count = round((count/nrow(dat))^2, 3),
                            Gini = round(DescTools::Gini(SATISFACTION_LEVEL),3),
                            Weighted_gini = round(Gini * norm_mean * norm_count, 3))

knitr::kable(df_sex1, align="lrrrrrrr")
```


```{r}
df_sex2 <- subset(df_sex1, select=c("SEX","Weighted_gini"))
df_sex3 <- subset(df.sex_decomp, df.sex_decomp$Categories %in% c("gini_between","gini_overlap"))
colnames(df_sex3) <- c("SEX","Weighted_gini")

df_sex <- rbind(df_sex2, df_sex3)
df_sex$Proportion <- round((df_sex$Weighted_gini / sum(df_sex$Weighted_gini)) * 100, 3)
rownames(df_sex) <- NULL

knitr::kable(df_sex, align="lrr")
```

```{r}
items <- c(rev(c("Male","Female")), "gini_between","gini_overlap")
color_items <- rev(c("#e41a1c","#377eb8","#a65628","#f781bf"))
lab_legend <- rev(c("Overlapping (45.294%)",
                    "Between groups (3.529%)", 
                    "Within Male (33.529%)",
                    "Within Female (17.647%)"))

# FACTORIZE ---
df_sex$SEX <- factor(df_sex$SEX, levels=items)
df_sex$y <- "Sex"

# PLOT ---
p1 <- ggplot(data=df_sex, aes(y=y, x=Proportion)) +
        geom_bar(aes(fill=SEX), stat="identity", color="#000000") +
        scale_fill_manual(values=color_items, labels=lab_legend) +   
        theme(
          axis.text.x = element_text(size = 10, color="#000000"),
          axis.text.y = element_blank(),
          axis.title.x = element_text(size = 10, color="#000000"),
          axis.title.y = element_blank(),
          plot.title = element_text(size = 10, color="#000000", hjust=0.5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks.x = element_line(size=0.4, color="#000000"),	
          axis.ticks.y = element_blank(),
          strip.text = element_text(size=10, color="#000000"),
          strip.background = element_rect(fill="#FFFFFF", color="#FFFFFF"),
          panel.background = element_rect(fill="#FFFFFF", color="#000000"),
          legend.text = element_text(size = 10, color="#000000"),
          legend.title = element_blank(),
          legend.key.size = unit(0.5, "cm"),
          legend.position = "bottom") +
        guides(fill=guide_legend(nrow=2)) +
        ylab("") +
        xlab("Proportion") + 
        ggtitle("")
```

```{r fig.height = 4, fig.width = 12, fig.align = "center"}
p1
```


</br></br>

#### *By Ethnicity*
```{r}
df_eth1 <- dplyr::group_by(dat, ETHNICITY) %>%
                  dplyr::summarise(count = n(),
                            mean = round(mean(SATISFACTION_LEVEL, na.rm = TRUE), 3),
                            sd = round(sd(SATISFACTION_LEVEL, na.rm = TRUE), 3),
                            norm_mean = round(mean/mean_satisfaction_level, 3),
                            norm_count = round((count/nrow(dat))^2, 3),
                            Gini = round(DescTools::Gini(SATISFACTION_LEVEL),3),
                            Weighted_gini = round(Gini * norm_mean * norm_count, 3))

knitr::kable(df_eth1, align="lrrrrrrr")
```


```{r}
df_eth2 <- subset(df_eth1, select=c("ETHNICITY","Weighted_gini"))
df_eth3 <- subset(df.eth_decomp, df.eth_decomp$Categories %in% c("gini_between","gini_overlap"))
colnames(df_eth3) <- c("ETHNICITY","Weighted_gini")

df_eth <- rbind(df_eth2, df_eth3)
df_eth$Proportion <- round((df_eth$Weighted_gini / sum(df_eth$Weighted_gini)) * 100, 3)
rownames(df_eth) <- NULL

knitr::kable(df_eth, align="lrr")
```



```{r}
items <- c(rev(c("Bahun","Chhetri","Dalit-Others","Gurung-Magar","Newar","Tamang")), "gini_between","gini_overlap")
color_items <- rev(c("#e41a1c","#377eb8","#4daf4a","#984ea3","#ff7f00","#ffff33","#a65628","#f781bf"))     
lab_legend <- rev(c("Overlapping (54.118%)",
                    "Between groups (25.294%)", 
                    "Within Bahun (1.176%)",
                    "Within Chhetri (1.765%)",
                    "Within Dalit-Other (1.765%)",
                    "Within Gurung-Magar (4.706%)",
                    "Within Newar (0.588%)",
                    "Within Tamang (10.588%)"))

# FACTORIZE ---
df_eth$ETHNICITY <- factor(df_eth$ETHNICITY, levels=items)
df_eth$y <- "Ethnicity"
        
# PLOT ---
p2 <- ggplot(data=df_eth, aes(y=y, x=Proportion)) +
        geom_bar(aes(fill=ETHNICITY), stat="identity", color="#000000") +
        scale_fill_manual(values=color_items, labels=lab_legend) +   
        theme(
          axis.text.x = element_text(size = 10, color="#000000"),
          axis.text.y = element_blank(),
          axis.title.x = element_text(size = 10, color="#000000"),
          axis.title.y = element_blank(),
          plot.title = element_text(size = 10, color="#000000", hjust=0.5),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks.x = element_line(size=0.4, color="#000000"),	
          axis.ticks.y = element_blank(),
          strip.text = element_text(size=10, color="#000000"),
          strip.background = element_rect(fill="#FFFFFF", color="#FFFFFF"),
          panel.background = element_rect(fill="#FFFFFF", color="#000000"),
          legend.text = element_text(size = 10, color="#000000"),
          legend.title = element_blank(),
          legend.key.size = unit(0.5, "cm"),
          legend.position = "bottom") +
        guides(fill=guide_legend(nrow=4)) +
        ylab("") +
        xlab("Proportion") + 
        ggtitle("")
```

```{r fig.height = 4, fig.width = 12, fig.align = "center"}
p2
```





</br></br>

#### *By Income Group*
```{r}
df_inc <- dplyr::group_by(dat, INCOME_LEVEL) %>%
            dplyr::summarise(count=n(), Gini=round(DescTools::Gini(SATISFACTION_LEVEL),3))

df_inc1 <- dplyr::group_by(dat, INCOME_LEVEL) %>%
                  dplyr::summarise(count = n(),
                            mean = round(mean(SATISFACTION_LEVEL, na.rm = TRUE), 3),
                            sd = round(sd(SATISFACTION_LEVEL, na.rm = TRUE), 3),
                            norm_mean = round(mean/mean_satisfaction_level, 3),
                            norm_count = round((count/nrow(dat))^2, 3),
                            Gini = round(DescTools::Gini(SATISFACTION_LEVEL),3),
                            Weighted_gini = round(Gini * norm_mean * norm_count, 3))

knitr::kable(df_inc1, align="lrrrrrrr")
```


```{r}
df_inc2 <- subset(df_inc1, select=c("INCOME_LEVEL","Weighted_gini"))
df_inc3 <- subset(df.inc_decomp, df.inc_decomp$Categories %in% c("gini_between","gini_overlap"))
colnames(df_inc3) <- c("INCOME_LEVEL","Weighted_gini")

df_inc <- rbind(df_inc2, df_inc3)
df_inc$Proportion <- round((df_inc$Weighted_gini / sum(df_inc$Weighted_gini)) * 100, 3)
rownames(df_inc) <- NULL

knitr::kable(df_inc, align="lrr")
```



</br></br>

### Calculate Correlation between mean happiness and standard deviation
```{r}
shapiro.test(log(df_eth1$mean))
shapiro.test(log(df_eth1$sd))
cor.test(log(df_eth1$mean), log(df_eth1$sd), method="pearson")
```


### Decomposition of standard deviation (varaince) with ANOVA
```{r}
aov_sex <- aov(SATISFACTION_LEVEL ~ SEX, data = dat)
summary(aov_sex)
```

```{r}
aov_ethnicity <- aov(SATISFACTION_LEVEL ~ ETHNICITY, data = dat)
summary(aov_ethnicity)
```


### Correlation between mean happiness and Gini by Ethnicity
```{r}
shapiro.test(log(df_eth1$Gini))
cor.test(log(df_eth1$mean), log(df_eth1$Gini), method="pearson")
```
